
## Uncertainties

```{r niruncer,results='hide',warning=FALSE,message=FALSE,echo=FALSE}
tabs("niruncert",paste0("Sector Agriculture: EU-28 uncertainty estimates"))


```

`r tabs("niruncert",display="cite")` shows the total EU-28 uncertainty estimates for the sector Agriculture and the uncertainty estimates for the relevant gases of each source category. The highest level uncertainty was estimated for N2O from 3D and the lowest for CH4 from sector 3A. With regard to the uncertainty on trend N2O from sector 3J shows the highest uncertainty estimates, CH4 from sector 3A the lowest. For a description of the Tier 1 uncertainty analysis carried out for the EU-28 see Chapter 1.6.

`r tabs("niruncert",display="full")` 

**Table to be added in May submissions**

Note: Emissions are in Gg CO2 equivalents; trend uncertainty is presented as percentage points; the sum of the source category emissions may not be the total sector emissions of the EU-NIR because uncertainty estimates are not available for all source categories in each of this 28 EU countries



## Sector-specific quality assurance and quality control and verification

### Introduction

This section gives an overview of the QA/QC procedures applied specifically for the agriculture sector of the EU GHG inventory.
It first gives an overview of the development of the agriculture QA/QC system with an outlook of further improvements to be discussed and/or implemented in coming years. A brief description of the QA/QC procedures used to process the data and interact with the countries is given.

This is followed by brief summaries of selected activities that have been carried out in the past to improve and/or verify national and EU wide GHG emissions from agriculture in the frame of the EU GHG inventory system. The list is not comprehensive.


### Improvements

#### Brief overview of the development of the QA/QC in the agriculture sector

A major revision of the present chapter on methodological issues and uncertainty in the sector agriculture was done for the submission in 2006 giving for the first time a complete overview of all relevant parameters required for the estimation of GHG emissions and the calculation of all background parameter in the CRF tables for agriculture.

The changes were partly due to a natural evolution of the inventory generation over the years and partly motivated by recommendations made by the UNFCCC review team on the occasion of the in-country review in 2005. The main issues raised by the Expert Review Team in 2005 and the major changes included (i) more transparent overview tables on methodological issues; (ii) better presentation of trend development; (iii) streamlining information contained in CRF and NIR; (iv) continuous working with countries in order to improve the inventory and allowing the quantification of all background data; (v) including a summary of workshops. For the submission in 2007, several errors identified in the background tables of the countries could be eliminated, thus improving the calculation of EU-wide background information. Further details were added to the inventory report for the submission in 2008, based on recommendations by the Expert Review Team of the in-country review in 2007. For the submissions in 2009 through 2014, background information was further developed.

In 2008, a novel approach to calculate uncertainties at the EU level including the assessment of the quality of the emission estimates at MS and EU level has been implemented and described in the NIR. This method was presented during the in-country-review in 2007 and its implementation in the EC-IR was suggested by the ERT. This has been complemented by a series of tables giving background information for the estimates of the uncertainty levels for activity data and emission factors.

Over the time, several sections were added describing specific QA/QC and verification activities (see also sections below), such as:

* Summary of the workshop on &apos;Inventories and Projections of Greenhouse Gas Emissions from Agriculture&apos; (2003)
* Summary of the findings of the GGELS project (Evaluation of the livestock sector&apos;s contribution to the EU greenhouse gas emissions. 
* A comparison between submissions and data from the FAO GHG database (2014)
* An analysis on the share of manure excretion by IPCC climate zones with EU wide independent data
* A description of the Survey on agricultural production methods (SAPM 2010)
* A summary of the LiveDate project on Nitrogen Excretion factors
* Workshop on improving national inventories for agriculture (2013)
* Comparison of Cultivated Organic Soil at the FAO GHG database and JRC calculations



#### Major changes for the 2015 submissions
The  submission in the year 2015 the QA/QC system brought a complete revision of the approach taken for the EU GHG inventory report in general and for the agriculture chapter in particular, driven by the need to adapt to new CRF software, increased number of countries to describe, and a series of new communication software products (e.g. EEA review tool, EU-GIRP). For this purpose, the EU GHG inventory was thoroughly revised. While this was true for the whole EU GHG inventory, this was particularly true for the agriculture sector. The following specific issues with regard to the GHG inventory in the agriculture sector were identified to require improvements:

* Focusing of the agriculture chapter in the EU-GHG inventory report on key categories and factors and parameters which have a significant relevance for EU total emissions. 

* The agricultural chapter applied a specific methodology to calculate "Tier levels" and aggregated uncertainties to more accurately account for correlation between the uncertainty estimates of the individual countries. The methodology was developed for the EU GHG inventory and published in peer-reviewed literature[^leip2010]. While this method was shown to provide additional insight for the uncertainty assessment of the EU GHG inventory, it was of no practical relevance for the overall GHG inventory, as a different method was used for other sectors. It was therefore decided to be not continued.

* One major drawback of previous GHG inventories was the difficulties to account for 'other' animal types or nitrogen inputs. With the new data processing framework[^eugirp], *all* data are now available so that a comprehensive analysis is possible

* Streamlining with other sector chapters was improved, not the least by using of harmonized plots to present trend-data at EU level while also showing data from those countries contributing most to EU values

* Addition of further sector-specific checks, in cooperation with the team working on the review of the submissions under the Emissions Sharing Directive. These checks focus on the consistency of reporting within the agricultural sector, the completeness of the background information provided which are important for transparence, in addition to issues related to correct implementation of recommended methodologies.

* The writing of the agriculture chapter of the EU-GHG inventory report has been highly automated[^prezi]. The process is directly based on the data submitted by the countries and are calculated on the fly thus no quantitative data are introduced &apos;manually&apos;. This allows to provide a report with quantitative information avoiding inconsistencies with the CRF data.

The newly developed system is described in the section QA/QC system in the agriculture sector.

[^leip2010]: &#09;Leip, A., 2010. Quantitative quality assessment of the greenhouse gas inventory for agriculture in Europe. Clim. Change. 103, 245-261. doi:http://dx.doi.org/10.1007/s10584-010-9915-5.
[^eugirp]: &#09;EU-GIRP: EU-Greenhouse gas Inventory Reporting and Plots, see https://github.com/aleip/eealocatorplots.git
[^prezi]: &#09;For an overview of the QA/QC system of the agriculture sector for the 2013 GHG inventory see presentation given for the ICR2013 at https://prezi.com/f1d3elxzd4qn/20131002_icr_agri/

#### Main improvements since 2019
Since the 2019 submission, a new check has been introduced on the correlation between milk yield versus feed digestibility and nitrogen excretion rates. Several parts of the code has been revised to align with current programming libraries the system implemented in 2015 was further developed.
***@Janka and @Alexander, please check if you remember other new elements introduced this year OR what we should do because of recommnendations and what we maybe haven't done yet???***


#### Further improvements

The following further improvements are foreseen for the next submission: 

* Further addition of sector-specific checks as required
* Further development of the comparison with FAO and CAPRI data

### QA/QC system in the agriculture sector

#### Quality checks

Several quality checks are performed in the EU-GIRP[^eugirp] software. 
They are documented in various modules of EU-GIRP and can be examined in the open source repository.
The checks include:

`r ## Update annually number of ERT recommendations`

* **Recommendations**: Country were checked if they had implemented last years' recommendations from the ESD review and from the UNFCCC review. From all recommendations, 16 were still unresolved and therefore the corresponding issues were reopened. ***@Janka please update this paragraph***

* **Check on NEs**[^checknes] and empty cells has been done by extracting all reported &apos;NE&apos;s and the empty cells, respectively, from the data base.  

* **Notation keys**: we identified emission categories where a country reported a notation key, while 20 ***@Janka please update this number*** or more countries reported emission estimates, in order to assess the potential over/underestimations (these also contained in NE checks in the file above). 

* **Outliers in activity data and emissions**: Data were checked on outliers in AD and emissions. For each source category the share of AD and emissions by the countries to total `r eukp` values were determined. A share above 95% was further assessed and in case this was not linked to a source category which is dominated by single countries (such as emissions from buffalo, which are dominated by Italy) the country was notified 

* **Check on erroneous units**: In several case, countries report background data using different units (e.g. fractions instead of percent values or vice versa; values per day instead of per year of vice versa; absolute values instead of values per head etc.). While these inconsistencies do not influence the reported emission estimates, a harmonization (at `r eukp` level) is important to ensure correct comparison of countries' values and a correct calculation of `r eukp` background data. An automated check[^unitcheck] is carried out detecting seven cases which can easily be recognised. Other &apos;mistakes&apos; in units used were detected following the outlier analysis (see below). The countries were notified via the review tool. 

* **Within-country outliers**: within-country outliers in IEFs and other parameters are detected on the basis of the distribution of the values provided[^outlier]. We used the method based on the mean values and the standard deviation. Specifically, those values were identified as outliers which were more distant from than 1.5 time the standard deviation in the data from the mean (both in positive and negative direction). As an additional criterion, the relation to the median was used. In case the value was within 10% of the median it was not considered as an outlier. This removed cases where a country uses a country-specific parameter while most countries use the default value.

* **Identification of potentially significant issues**: For each of the outliers identified it was determined whether or not this could be a potentially significant issue based on the criterion of a share of 0.05% of national total GHG emissions. The 'size' of the possible over- or under-estimation was quantified comparing the reported value with an estimate using the median IEF or parameter as reported by all countries[^potsig]. All outliers were &apos;manually&apos; cross-checked and analysed. Countries were notified on the results of the analysis.

* **Time series outliers/inconsistencies**: Time series outliers were detected on the basis of the same method as also used for the within-country-outlier check. Basis for the underlying distribution of data in this case, however, was not the values reported from all countries during the whole time series, but only the data reported by the country assessed. Only growth rates larger than &#177;3% qualify as 'outliers'. However, this generates a large number of potential outliers which require further assessment. The following types of &apos;issues&apos; were identified, which might be linked either to an inconsistent time series or be the consequence of 'real' trends:
* * *Period outphased*:	Relative constant trend with few years above/below the trend that &apos;looks plausible&apos;.
* * *Trend break*: Time series in steps, in a stair shape: a few similar values, then a jump, and the same again.
* * *One break group trend*: Regular time series with a different trend for a group of years, and a step when jumping from/coming back to the general trend.
* * *Inflection point*: Trend suddenly changes from a specific year from which the growth of the values changes sign.
* * *Single outlier*: One or few isolated year(s) where the value is out of the general trend
* * *Smooth group trend change*: A series of years where the trend changes compared to the rest of the time series, but without any jumps
* * *Trend jump*: There is a jump at some point in the time trend but it continues running parallel to the first section, after the jump. 
* * *Jump and shape*: There is a jump at some point in the time trend and, after the jump, the trend changes shape

Time series outliers and inconsistencies can in many cases be explained by socio-economic fluctuations, crises, or other events influencing the time series of populiations or emission factors. Explanations received by countries are collected and the issue flagged representing a real trend.

* **Sector-specific checks**: Several checks were performed tailored to the reporting in the sector agriculture[^adchecks]^,^[^nexchecks]. First, the data are checked on consistency in reporting of activity data throughout the tables. Further, several other tests are performed:
* * Difference between the sum of nitrogen excreted and reported in the different manure management system (MMS) versus the total reported nitrogen excreted
* * Difference between the total nitrogen excreted and the product of animal population and nitrogen excretion rate
* * Difference of the sum of N handled in MMS over animal type vs. total N handled in each MMS
* * Check of the reported IEF per MMS with the total N excreted and the reported emissions
* * Calculation and evaluation of the IEF in category 3.B.2 by animal type and in relation to the total N excreted
* * Check that the sum of manure allocated to climate regions adds up to 100% over all MMS and climate regions
* * Check that compares the Manure 'managed' in Pasture Range and Paddock in category 3.B.2 with AD in 3.D.1.3 (Urine and Dung Deposited by Grazing Animals). The sum of FRPR over all animal types should therefore equal the AD in category 3.D.1.3.
* * Comparison of the IEF in 3.D.1.3 (N2O emissions from Urine and Dung Deposited by Grazing Animals) with default IEFs EF3_RPR_CPP for Cattle - Pigs and Poultry (0.02) and, EF3_RPR_SO for Sheep and other animals (0.01) using the shares FracRPR_CPP and FracRPR_SO of manure deposited by the two animal groups.
* * Comparison of the fraction of N lost in MMS (via volatilization of NH3+NOx) versus total managed manure. According to IPCC Table 10.22 most of the loss fractions are between 20% and 45% of N in managed manure and N loss ratios are identified that are higher than 45% or lower than 20%.
* * Comparison of the manure 'managed' and not lost as NH3+NOx or leaching in MMS (3B2) with Animal manure applied to soil (3D12a). Manure available for application is obtained from N managed in MMS and not lost (FracLOSSMS) according to Table 10.23 plus any addition of bedding material. The loss fractions in Table 10.23 include also losses of N2, which are not included in the indirect emissions-volatilisation. Therefore, FAM is expected to be smaller than N managed in MMS minus N lost as NH3+NOx+leaching unless bedding material has been accounted for. In case of crop residues as bedding material care has to be taken to avoid double counting.

* **Recalculation**: Countries were asked for justifications of recalculations of more than 0.05% of national total emissions (excluding LULUCF) for years `r firstyear` and `r lastyear`, focusing on key categories.  


```{r checkstats,results='hide',warning=FALSE,message=FALSE,echo=FALSE}
checkstats<-as.data.frame(matrix(rep(NA,25),ncol=5,nrow=5))
names(checkstats)<-c("Check","Resolved","Partially resolved","Unresolved","Not yet responded")
checkstats$Check<-c("Completeness","Outliers","Recalculations","Time series","ESD review")
complete<-c("50%","19%","31%","0%")
outliers<-c("24%","2%","8%","65%")
recalcul<-c("90%","0%","10%","0%")
timeseri<-c("55%","3%","22%","21%")
esdrevie<-c("33%","17%","33%","17%")

response<-rbind(complete,outliers,recalcul,timeseri,esdrevie)
checkstats[,2:5]<-response
checkdate<-"March 15, 2019"
tabs("checkstat",paste0("Status of issues as of ",checkdate))

```
***@Alexander please update text/table below until end of section*** A much lower number of issues were identified, compared to last year (340 issues were identified in 2017):

* 12 completeness issues (related to ‘NE’/’empty’/’notation keys’)
* 11 country-outlier issues
* 14 recalculation issues
* 16 time trend issues
* 16 recommendations
* 34 agricheck issues
* 17 other issues (wrong units, same values reported for 2015 and 2016)

`r #In 2017 agrichecks were done only in March due to time constraints. Some of the countries responded to the issues but they did not have to, given that questions were sent when it was not possible to submit new data. In this situation, none of these issues was triggered to the second step review, but we just kept record to check the issues again next year (or flag them as solved, if it is the case).`

The status of responses as of `r checkdate` is given in `r tabs("checkstat",display="cite")`:

`r tabs("checkstat",display="full")`

`r pander(checkstats)`&nbsp;

All issues had been responded by `r checkdate`, being the time series the type of issues with the highest percentage of resolution, followed by the recalculations. Most of the time series and the recalculation issues required just an explanation and some of the recalculations were due to mistakes which have mostly been corrected. The agrichecks are the type of issues with the highest share of questions still unresolved, often requiring detailed information which sometimes the countries cannot easily obtain. Similarly happens with the recommendations, 63% of which have not been resolved yet but countries are working on getting the necessary data for the resolution of the issues. Regarding outliers, 63% of the issues are still partly resolved or unresolved, probably needing more time to justify.


#### Calculation of EU background data 


EU-wide background data were calculated as weighted averages of the parameters provided by the countries, using activity data (animal numbers in category 3A and 3B and N input in category 3D) as weighting factors[^wave].

Care is being taken to not include in the calculation erroneous values:

* Data which had been identified as being reported with a different unit than the values reported by other countries (see above) were *converted* into the appropriate unit before calculating `r eukp` weighted averages

* Data which *obviously* wrong (very large outliers) but for which no clear correction could be identified were *eliminated* from the calculation of the `r eukp` weighted averages to avoid biases in the results. Therefore, the `r eukp` weighted averages - in some cases - could not represent 100% of `r eukp` activity data.

[^wave]: &#09;https://github.com/aleip/eealocatorplots/blob/master/eugirp_euweightedaverages.r


#### Compilation of the chapter agriculture for the EU-GHG inventory report

The agriculture chapter of the EU-GHG inventory report takes advantage of the data base generated by EU-GIRP. All numeric data presented in the chapter are calculated directly using the processed data as described above, thus eliminating the risk of transcription or copy errors. This does not eliminate the possibility of mistakes completely. Therefore, all values are cross-checked.


[^checknes]: &#09;https://github.com/aleip/eealocatorplots/blob/master/eugirp_checknes.r
[^unitcheck]: &#09;https://github.com/aleip/eealocatorplots/blob/master/eugirp_checkunits.r
[^outlier]: &#09;https://github.com/aleip/eealocatorplots/blob/master/eugirp_checkoutliers.r
[^potsig]: &#09;See function `ispotentialissue()` in the file https://github.com/aleip/eealocatorplots/blob/master/eugirp_functions.r
[^adchecks]: &#09;https://github.com/aleip/eealocatorplots/blob/master/agrichecks1ADs.r
[^nexchecks]: &#09;https://github.com/aleip/eealocatorplots/blob/master/agrichecks2Nex.r



